Enhancing Student Engagement via Reduction of Frustration with Programming Assignments using Machine Learning

### 2017-04-29 paper.tex
Retrieved the figure back from original version
And some paragraphs from related work. Also reformatted figures to fit
them in the page. Closes #3.

@@

@@

\begin{figure*}[h!t] 
\centering 
\includegraphics[width=3in]{editorRresult.png} 
\caption{Web-based editor used for code evaluation.}
\label{fig_editor} 
\end{figure*}

\begin{figure*}[hb!] 
\centering 
\includegraphics[width=3in]{KeyDyn.png} 
\caption{Keystroke Dynamics, adapted from Sim \& Janakiraman \cite{sim2007digraphs}.}
\label{fig_KD} 
\end{figure*}

@@

\section{Experiment and results}
\label{sec:exp}
The aim of this research is to evaluate the use of keyboard
and mouse dynamics as an appropriate sensory input for an affective recognition
system. The context of use is a learning environment for programmers. In
particular for learning activities consisting in writing short programs
interactively. An experimental approach was adopted with this aim: Sensory and
quantitative data was collected from learners as they were enrolled in a basic
course of Python programming. This data was then pre-processed using the method
described earlier in Section \ref{sec:method}, and together with the quantitative data obtained from users,
classifiers were trained and validated. The results were then compared
and we arrived to some conclusions regarding the effectiveness of the method. The goal given to
subjects was to solve as many exercises as they could in a period of two weeks.
There was neither time limit nor a minimum amount of time required for a
participant while trying to solve the exercises or complete the tutorial. The
participants were able to stop and resume their interaction with the system at
any time.
%
\begin{figure*}[!t] 
\centering 
\includegraphics{Completed.png} 
\includegraphics{classDist.png} 
\caption{Completed activities and exercises and by learner in ascending order (left) and lass distribution of emotions as reported by learners (right).}
\label{fig_completed} 
\end{figure*}

>>>>

Elliott, Rickel y Lester \cite{elliott1999lifelike,d2008autotutor} propose the integration of an affective reasoner
with an agent in a virtual environment. In this work an agent called Steve
responds emotionally to other agents and interactive users. The agent simulates
his emotions through different multimedia modes including facial expressions and
speech. Affective reasoning agents were used for training, by putting students
in work related situations. Agents could react to the behavior of students, for
instance if a student was being careless in a task and was in a dangerous
situation, Steve would show distress or fear. Also in a virtual environment the
work of McQuiggan, Robison \& Lester \cite{mcquiggan2010affective} extends this line of research by
investigating the affective transitions that occur throughout narrative-centered
learning experiences. The analysis of affective state transitions in this work
replicated the findings by D’Mello et al. \cite{d2008autotutor} and Baker et al. 
\cite{rodrigo2009affective} where
also engagement/flow dominated self-reported affect. For outdoor environments
Shen, Wang \& Shen \cite{shen2009affective} augment a pervasive e-learning platform with affective
recognition.  The results about emotion recognition from physiological signals
achieved a best-case accuracy (86.3\%) for four types of learning emotions.

@@

They also stress that their methods must be validated further in
other contexts. Moods may have an impact on programmer’s performance according
to Khan, Hierons y Brinkman \cite{khan2007mood}. It may be possible to detect moods
on the basis of information regarding the programmer’s use of the keyboard and
mouse, and to integrate them into development environments that can improve
programmer performance. They briefly describe a further experiment that could
use keyboard and mouse but only as future work. There are other studies about
the behaviour of programmers, which are not directly concerned with affective
states but are nevertheless important to their performance. Eye tracking in
computing education is proposed in the work Busjahn et al., \cite{busjahn2014eye}, and it is
also used for assessing learner´s comprehension of C++ and Python programs in
Turner et al., \cite{turner2014eye}. Blikstein \cite{blikstein2011using} proposed 
an automated technique to
assess, analyse and visualize students learning computer programming. Blikstein
employs different quantitative techniques to extract students’ behaviours and
categorize them in terms of programming experience.

\begin{figure*}[h!tbp] 
\centering 
\includegraphics[width=0.9\textwidth]{KMDAffective.png} 
\caption{Code Execution and Affect Recognition Architecture.}
\label{fig_process} 
\end{figure*}

@@

\begin{figure*}[h!t] 
\centering 
\includegraphics[width=0.9\textwidth]{editorRresult.png} 
\caption{Web-based editor used for code evaluation.}
\label{fig_editor} 
\end{figure*}

\begin{figure*}[hb!] 
\centering 
\includegraphics[width=0.9\textwidth]{KeyDyn.png} 
\caption{Keystroke Dynamics, adapted from Sim \& Janakiraman \cite{sim2007digraphs}.}
\label{fig_KD} 
\end{figure*}

@@

\section{Experiment and results}
\label{sec:exp}
The aim of this research is to evaluate the use of keyboard
and mouse dynamics as an appropriate sensory input for an affective recognition
system. The context of use is a learning environment for programmers. In
particular for learning activities consisting in writing short programs
interactively. An experimental approach was adopted with this aim: Sensory and
quantitative data was collected from learners as they were enrolled in a basic
course of Python programming. This data was then pre-processed using the method
described earlier in Section \ref{sec:method}, and together with the quantitative data obtained from users,
classifiers were trained and validated. The results were then compared
and we arrived to some conclusions regarding the effectiveness of the method. The goal given to
subjects was to solve as many exercises as they could in a period of two weeks.
There was neither time limit nor a minimum amount of time required for a
participant while trying to solve the exercises or complete the tutorial. The
participants were able to stop and resume their interaction with the system at
any time.
%
\begin{figure*}[!t] 
\centering 
\includegraphics[width=0.9\textwidth]{Completed.png} 
\includegraphics[width=0.9\textwidth]{classDist.png} 
\caption{Completed activities and exercises and by learner in ascending order (left) and lass distribution of emotions as reported by learners (right).}
\label{fig_completed} 
\end{figure*}
